### 2014-11-04 alternativefigure3.rmd
fiks

---
title: "Alternative Overview Figures"
author: "Vegard"
date: "Thursday, October 30, 2014"
output: html_document
---

Two scenarios

- Two groups, three batches. False Positives due to batch and ComBat.
- Three groups, two batches. False negatives due to batch, false positve due to mean adjust.

### False Positives due to batch and ComBat.

lage data set, 1000gener, 10+10
ingen gruppe effect eller batch effect.

1. plotte, målineg + meanstrek + CI + FC + p

2.
Dele i batcher 9+2+9, legge til batch effect.
plotte, målineg + meanstrek + CI + FC + p med batch separator i plottet.

3. Combat adjust
plotte, målineg + meanstrek + CI + FC + p uten batch separator i plottet.

4 plot med 2way anova og least square. Funker vel neppe med 1+1 bathen?

```{r}
library(sva)
source("commonscripts/helperfunctions.r")
source("commonscripts/boxplot_function_v3.r")
library(limma)
library(lsmeans)

set.seed(100)
ngenes = 1000
index=1
sa = createsampleannotation(  list(c(8,0), c(2,2), c(0,8)), as.factors=TRUE)
matrix_true = matrix(rnorm(ngenes * nrow(sa), mean=5, sd=1), nrow=ngenes, ncol=nrow(sa))

```

```{r, dev="png"}
plot_one_gene(matrix_true[index,], group=sa$group, main=paste("True values"),
              estimatemethod="CI")
```

Add batch effect.
```{r, dev="png"}
matrix_batcheffect = matrix_true
matrix_batcheffect[,sa$batch==1] = matrix_batcheffect[,sa$batch==1]+1
matrix_batcheffect[,sa$batch==3] = matrix_batcheffect[,sa$batch==3]-1
plot_one_gene(matrix_batcheffect[index,], group=sa$group, batch=sa$batch,
              main=paste("Batch affected values"), estimatemethod="CI")
```

Adjust for batch effect.
```{r, dev="png"}

mod0 = model.matrix(~1,data=sa)
mod = model.matrix(~as.factor(sa$group))

matrix_batchadjusted = ComBat(dat=matrix_batcheffect, batch=as.factor(sa$batch), 
                      mod=mod, numCovs=NULL, 
                      par.prior=TRUE, prior.plots=FALSE)
matrix_batchadjusted = removeBatchEffect(matrix_batcheffect, batch=as.factor(sa$batch),
                                         design=mod)

plot_one_gene(matrix_batchadjusted[index,], group=sa$group, batch=sa$batch,
              main=paste("Batch effect adjusted values"), estimatemethod="CI")
```

```{r, dev="png"}
plot_one_gene(matrix_batcheffect[index,], group=sa$group, batch=sa$batch,
              main=paste("Batch affected values, batch included in ANOVA"),estimatemethod="lsmeans")

```

```{r}
#batch2FC = mean(matrix_true[index,sa$group==1 & sa$batch==2]) - 
#          mean(matrix_true[index,sa$group==2 & sa$batch==2])

figfilename = file.path( getwd(), "plots", paste("boxplots_v3_anovaadjusted", sep=""))
#figfile = paste( figfilename, ".png", sep=""); png(file = figfile, width=1600, height=800)
figfile = paste( figfilename, ".pdf", sep=""); 
pdf(file =figfile, width=6, height=6)
op=par(mfrow=c(4, 1),   xaxt="n", mar=c(2,3,2,1)+0.1)
ylim=c(min(matrix_batcheffect[index,]), max(matrix_batcheffect[index,]))
plot_one_gene(matrix_true[index,], group=sa$group, main=paste("(a) True values"),
              estimatemethod="CI", ylim=ylim)
plot_one_gene(matrix_batcheffect[index,], group=sa$group, batch=sa$batch,
              main=paste("(b) Batch affected values"), estimatemethod="CI", ylim=ylim)
plot_one_gene(matrix_batchadjusted[index,], group=sa$group, batch=sa$batch,
              main=paste("(c) Batch effect adjusted values"), estimatemethod="CI", ylim=ylim)
plot_one_gene(matrix_batcheffect[index,], group=sa$group, batch=sa$batch, ylim=ylim,
              main=paste("(d) Batch affected values, batch included in ANOVA"),estimatemethod="lsmeans")

dev.off()

```

### False Positives due to batch and mean adjustment.

```{r , dev="png"}

#set.seed(100)
ngenes = 1000
index=3
sa = createsampleannotation(  list(c(8,3,0), c(0,7,8)), as.factors=TRUE)
matrix_true = matrix(rnorm(ngenes * nrow(sa), mean=5, sd=1), nrow=ngenes, ncol=nrow(sa))

# add a group effect for group 3
matrix_true[,sa$group==3] = matrix_true[,sa$group==3]+4
plot_one_gene(matrix_true[index,], group=sa$group, main=paste("True values"),
              estimatemethod="CI")

```

Add batch effect.
```{r , dev="png"}
matrix_batcheffect = matrix_true
matrix_batcheffect[,sa$batch==1] = matrix_batcheffect[,sa$batch==1]-2
matrix_batcheffect[,sa$batch==2] = matrix_batcheffect[,sa$batch==2]+2
plot_one_gene(matrix_batcheffect[index,], group=sa$group, batch=sa$batch,
              main=paste("Batch affected values"), estimatemethod="CI")
```

Batch adjust with mean center.

```{r , dev="png"}
matrix_batchadjusted = matrix_batcheffect
genemeans = rowMeans(matrix_batcheffect)
matrix_batchadjusted[,sa$batch ==1] =  genemeans +
  matrix_batchadjusted[,sa$batch ==1] - rowMeans( matrix_batchadjusted[,sa$batch ==1])
matrix_batchadjusted[,sa$batch ==2] =  genemeans + 
  matrix_batchadjusted[,sa$batch ==2] - rowMeans( matrix_batchadjusted[,sa$batch ==2])

#index=index+1
plot_one_gene(matrix_batchadjusted[index,], group=sa$group, batch=sa$batch,
              main=paste("Batch effect adjusted values"), estimatemethod="CI")
```

Include batch as factor in ANOVA.
```{r , dev="png"}
plot_one_gene(matrix_batcheffect[index,], group=sa$group, batch=sa$batch,
              main=paste("Batch affected values, batch included in ANOVA"),estimatemethod="lsmeans")

```

```{r}

figfilename = file.path( getwd(), "plots", paste("boxplots_v3_meanadjusted" , sep=""))
figfile = paste( figfilename, ".pdf", sep=""); 
pdf(file =figfile, width=6, height=6)
ylim=c(min(matrix_batcheffect[index,]), max(matrix_batcheffect[index,]))
op=par(mfrow=c(4, 1),   xaxt="n", mar=c(2,3,2,1)+0.1)
plot_one_gene(matrix_true[index,], group=sa$group, main=paste("(a) True values"),
              estimatemethod="CI", ylim=ylim)
plot_one_gene(matrix_batcheffect[index,], group=sa$group, batch=sa$batch,
              main=paste("(b) Batch affected values"), estimatemethod="CI", ylim=ylim)
plot_one_gene(matrix_batchadjusted[index,], group=sa$group, batch=sa$batch,
              main=paste("(c) Batch adjusted values"), estimatemethod="CI", ylim=ylim)
plot_one_gene(matrix_batcheffect[index,], group=sa$group, batch=sa$batch, ylim=ylim,
              main=paste("(d) Batch affected values, batch included in ANOVA"),estimatemethod="lsmeans")

dev.off()

```

```{r}
# debug
# rart?
#plot(matrix_true[,10]-matrix_true[,11], matrix_batchadjusted[,10]-matrix_batchadjusted[,11])

#plot(matrix_true[,10]-matrix_true[,11], rowMeans(matrix_batchadjusted[,1:10])-rowMeans(matrix_batchadjusted[,11:20]))

```